Resolvable Ambiguity

### 2019-07-07 paper/paper.tex
Updated the abstract.

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Text of abstract \ldots.
\end{abstract}

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
% Why?

Ever since the early 60s, the problem of deciding if a context-free
grammar is unambiguous or not has been known to be undecidable. As a
consequence, a large number of restricted grammars have been developed
to guarantee that a language definition is unambiguous. This
traditional view of only allowing unambiguous grammars has until
recently been taken for granted as the only truth: \emph{the} way of
how the syntax of a language must be defined. However, up front
unambiguous grammars can force language designers to make arbitrary
choice to disambiguate the language, although the natural choice may be to
postpone the decision to the programmer. Moreover, an important
objective of domain-specific language development is to construct
languages by composing and extending existing
languages. Unfortunately, the composition of restricted unambiguous
grammars easily results in ambiguous grammars, which again are
undecidable. In this paper, we depart from the traditional view of
unambiguous grammar design, and allow ambiguities to be delayed until
compile time, allowing the user to perform the disambiguation. A
natural decision problems follows: given a language definition, can a
user always disambiguate an ambiguous program?  We introduce and
formalize this fundamental problem---called the \emph{resolvable
  ambiguity problem}---and divide it into separate static and dynamic
resolvability problems. We prove soundness and completeness for a
restricted language in the static case, and soundness in the general
case for dynamic resolvability. The approach is evaluated through
three separate case studies, covering both a large
existing programming language, and the composability of
domain-specific languages.
%A traditional view when designing programming languages is that syntax
%definitions should be unambiguous: only unambiguous grammars are used
%to construct language specific parsers that either rejects a program,
%or accepts it and produces a unique parse tree. However,
\end{abstract}
